\chapter{References} \label{chap:concl}

%technical paper - authors,"title". journal, volume, year.

%chapter in book - authors, "chapter" in Book, edition. Editor's names, Ed. Local of publishing: Publisher, year. 

%website -
%J. Geralds, “Sega Ends Production of Dreamcast,” vnunet.com, para. 2, Jan. 31,2007. [Online]. Available: http://nli.vnunet.com/news/1116995. [Accessed Sept.12, 2007].

[1] J. Liang. "Fifteen Percent Of Ocean Floor Now Mapped," \textit{deeperblue.com}, Jun. 2019. [Online]. Available: https://www.deeperblue.com/fifteen-percent-of-ocean-floor-now-mapped/. [Accessed Oct. 28, 2019]
\label{r:deeperblue}

%thesis
[2] A. M. Bonito. “Acoustic system for ground truth underwater positioning in DEEC’s test tank,” Master’s thesis, Faculty of Engineering of the University of Porto, Porto, Jul. 23 2019.
\label{r:afonso-thesis}

%technical paper - authors,"title". journal, volume, year.
[3] H. P. Tan, R. Diamant, W. K. G. Seah, M. Waldmeyer. "A survey of techniques and challenges in underwater localization". \textit{Ocean Engineering}, vol. 38, pp. 1663-1676, 2011.
\label{r:survey-tech-chall}

%book
[4] F. B. Jensen, W. A. Kuperman, M. B. Porter, H. Schmidt."Fundamentals of Ocean Acoustics" in \textit{Computational Ocean Acoustics}, second edition. Springer Science \& Business Media, pp.3-17, 2011. 
\label{r:ocean-acoust}


%Article in Daily, Weekly, or Biweekly Newspaper or Magazine - [11] B. Metcalfe, “The numbers show how slowly the Internet runs today,” \it(Infoworld), 30 Sep., p. 34, 2006.
[5] M. Stojanovic, J. Preisig. "Underwater Acoustic Communication Channels: Propagation Models and Statistical Characterization", \textit{IEEE Communications Magazine}, vol.47, p.84-89, 2009.
\label{r:commchan}

%thesis
[6] J. M. F. Magalh\~{a}es. "Improving Time of Arrival Estimation Using Encoded Acoustic Signals," Master’s thesis, Faculty of Engineering of the University of Porto, Porto, Jul. 8 2018.
\label{r:thesis-joao}

[7] D. E. Chaitanya, C. V. Sridevi, G. S. B. Rao. "Path loss analysis of Underwater communication systems," \textit{IEEE Technology Students' Symposium}, pp. 65-70, 2011.
\label{r:pathloss}

[8] B. M. R. Bharathi, A. R. Mohanty. "Underwater Sound Source Localization by EMD-Based Maximum Likelihood Method," \textit{Acoustics Australia}, vol.46, no.2, pp. 193-203, 2018.
\label{r:crosscorr}

[9] X. You, Y. Wu, M. Zhu, X. Li, L. Zhang. "Low Complexity Short Baseline Localization Algorithm Based on Taylor Expansion," in 2019 IEEE International Conference on Signal Processing, Communications and Computing, 2019.
\label{r:sbl}

[10] K. G. Kebkala, A. I. Mashoshinc. "AUV Acoustic Positioning Methods," \textit{Gyroscopy and Navigation}, vol.8, no.1, pp. 80-89, Sep. 2017
\label{r:lblsblusbl}

[11] D. Sun, J. Ding, C. Zheng, W. Huang. "An Underwater Acoustic Positioning Algorithm for Compact Arrays With Arbitrary Configuration", in \textit{IEEE Journal of Selected Topics in Signal Processing}, vol. 13, no. 1, pp. 120-130, Mar. 2019
\label{r:usbl-new}

[12] EvoLogics GmbH. "Underwater USBL Positioning Systems," \textit{evologics.de}, 2018. [Online]. Available: https://evologics.de/usbl. [Accessed Jan. 26, 2020]
\label{r:evologics1}

[13] EvoLogics GmbH. "S2C Technology," \textit{evologics.de}, 2018. [Online]. Available: https://
evologics.de/s2c-technology [Accessed Jan. 26, 2020]
\label{r:evologics2}

[14] EvoLogics GmbH. "18/34 communication and positioning devices," \textit{evologics.de}, 2018. [Online]. Available: https://evologics.de/acoustic-modem/18-34 [Accessed Jan. 26, 2020]
\label{r:evologics3}

[15] EvoLogics GmbH. "S2C R 12/24 USBL communication and positioning device," \textit{evologics.de}, 2018. [Online]. Available: https://evologics.de/acoustic-modem/12-24/usbl-serie [Accessed Jan. 26, 2020]
\label{r:evologics4}

[16] Sonardyne. "MICRO-RANGER 2 SHALLOW WATER USBL SYSTEM," \textit{sonardyne.com}, 2020. [Online]. Available: https://www.sonardyne.com/product/micro-ranger-2-shallow-water-usbl-system/ [Accessed Feb. 3, 2020]
\label{r:sonardyne1}

[17] J.S. Esteves , A. Carvalho , C. Couto. "Generalized geometric triangulation algorithm for mobile robot absolute self-localization,"\textit{ 2003 IEEE International Symposium on Industrial Electronics}, Rio de Janeiro, Brazil, pp. 346-351, vol. 1, 2003
\label{r:triangalgo}

[18] A. P. Pandian, T. Senjyu, S. M. S. Islam, H. Wang. "Weighted Trilateratiton and Centroid Combined Indoor Localization in Wifi Based Sensor Network," in \textit{Proceeding of the International Conference on Computer Networks, Big Data and IoT (ICCBI - 2018)}, pp.163-166, Jul. 2019
\label{r:trilat}

[19] M. Erol-Kantarci, H. T. Mouftah, S. Oktug. "A Survey of Architectures and Localization Techniques for Underwater Acoustic Sensor Networks," in \textit{IEEE Communications Surveys \& Tutorials}, vol. 13, no. 3, pp. 487-502, 2011.
\label{r:arch_localiz}

[20] V. Chandrasekhar, W. KG Seah, Y. S. Choo, H. V. Ee. "Localization in underwater sensor networks: survey and challenges," \textit{WUWNet '06: Proceedings of the 1st ACM international workshop on Underwater networks}, pp. 33-40, Sep. 2006
\label{r:suvey-loc}

%book
[21] S. Gezici. "A Survey on Wireless Position Estimation," in \textit{Wireless Personal Communications}. Springer Science \& Business Media, vol.44, no.3, pp.263–282, Feb. 2008
\label{r:wirelesscomm}

[22] Kongsberg Maritime. "HiPAP - HIGH PRECISION ACOUSTIC POSITIONING"  \textit{kongsberg.com}, 2016. [Online]. Available: https://www.kongsberg.com/maritime/products/Acoustics-Positioning-and-Communication/acoustic-positioning-systems/hipap-models/ [Accessed Feb. 8, 2020]
\label{r:hipap_hardw}

